---
title: '**Lapse: Exploratory Data Analysis (EDA)**'
author: '*Roberto Jesús Alcaraz Molina*'
date: "04/05/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    df_print: paged
header-includes: 
- \usepackage{float}
- \usepackage{amsbsy}
- \usepackage{amsmath}
- \usepackage{graphicx}
- \usepackage{subfig}
- \usepackage{booktabs}
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      echo = T, 
                      message = FALSE,
                      fig.pos="H", 
                      fig.align="center",
                      fig.width=15,
                      cache=FALSE, error = TRUE)
```

```{r, echo = F}
# install.packages("pacman")
pacman::p_load(tidyverse, tidymodels, workflows, tune, patchwork, FactoMineR, dotwhisker, splines)
theme_set(theme_bw())

# Models packages
pacman::p_load(ranger)
```

```{r}
lapse_data <- readRDS("../00_data/insurance_lapse.RDS")
```

# Description and visualization of the variables

Our variables are:

- **Indicative variables**:

  - `ID`
  - `key`

- **Categorical variables (10 + 1)**:

  - `lapse` (target)
  - `period`
  - `cover`
  - `sex`
  - `smoker`
  - `good_health`
  - `capital_factor_1`
  - `capital_factor_2`
  - `capital_factor_3`
  - `IMC_factor_1`
  - `IMC_factor_2`


- **Numerical variables (7)**:

  - `actuarial_age`
  - `sus_age`
  - `exp`
  - `duration`
  - `duration_cat`
  - `capital`
  - `IMC`

## Target variable: `lapse`

```{r}
# lapse
lapse_data$lapse <- as.factor(lapse_data$lapse)
table(lapse_data$lapse)/nrow(lapse_data)
```

We are going to split the data in 70% for training and 30% for testing
```{r}
set.seed(123)

lapse_split <- initial_split(lapse_data, prop = 0.7, strata = lapse)
lapse_train <- training(lapse_split)
lapse_test  <- testing(lapse_split)
```

## Categorical variables (10 + 1)

- `period`
- `cover`
- `sex`
- `smoker`
- `good_health`
- `capital_factor_1`
- `capital_factor_2`
- `capital_factor_3`
- `IMC_factor_1`
- `IMC_factor_2`

### Period

```{r}
lapse_train %>%
  ggplot(aes(x = period, fill = lapse)) +
  geom_bar(position = "dodge") 

tab <- table(lapse_train$period, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

# In the third year the percentage increases
```

### Cover
```{r}
table(lapse_train$cover)/nrow(lapse_train)

lapse_train %>%
  ggplot(aes(x = cover, fill = lapse)) +
  geom_bar(position = "dodge")

tab <- table(lapse_train$cover, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

# as we increase the number of coverages, the proportion decreases
```

### sex
```{r}
table(lapse_train$sex)/nrow(lapse_train)

lapse_train %>%
  ggplot(aes(x = sex, fill = lapse)) +
  geom_bar(position = "dodge")

tab <- table(lapse_train$sex, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

# more or less equal, slightly higher in men
```

### smoker
```{r}
table(lapse_train$smoker)/nrow(lapse_train)

lapse_train %>%
  ggplot(aes(x = smoker, fill = lapse)) +
  geom_bar(position = "dodge")

tab <- table(lapse_train$smoker, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

# higher in non-smokers
```

### good_health 
```{r}
table(lapse_train$good_health)/nrow(lapse_train)
lapse_train %>%
  ggplot(aes(x = good_health, fill = lapse)) +
  geom_bar(position = "dodge")
tab <- table(lapse_train$good_health, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

# higher for the people with good health
```

### capital_DESC
```{r}
table(lapse_train$capital_factor_1)/nrow(lapse_train)
lapse_train %>%
  ggplot(aes(x = capital_factor_1, fill = lapse)) +
  geom_bar(position = "dodge")

tab <- table(lapse_train$capital_factor_1, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

table(lapse_train$capital_factor_2)/nrow(lapse_train)
lapse_train %>%
  ggplot(aes(x = capital_factor_2, fill = lapse)) +
  geom_bar(position = "dodge")
tab <- table(lapse_train$capital_factor_2, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

table(lapse_train$capital_factor_3)/nrow(lapse_train)
lapse_train %>%
  ggplot(aes(x = capital_factor_3, fill = lapse)) +
  geom_bar(position = "dodge")
tab <- table(lapse_train$capital_factor_3, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

# Non-linear?
```

### IMC_DESC

```{r}
lapse_train %>%
  ggplot(aes(x = IMC_factor_1, fill = lapse)) +
  geom_bar(position = "dodge")

tab <- table(lapse_train$IMC_factor_1, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

lapse_train %>%
  ggplot(aes(x = IMC_factor_2, fill = lapse)) +
  geom_bar(position = "dodge")

tab <- table(lapse_train$IMC_factor_2, lapse_train$lapse)
cbind(tab, tab[, 2] / (tab[, 1] + tab[, 2]))

# more or less the same
```

## Numerical variables (7)

- `actuarial_age`
- `sus_age`
- `exp`
- `duration`
- `duration_cat`
- `capital`
- `IMC`


### actuarial and suscription age
```{r}
lapse_train %>%
  ggplot(aes(x = actuarial_age, color=lapse, fill=lapse)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666") 
# seems good

# sus_age
lapse_train %>%
  ggplot(aes(x = sus_age, color=lapse, fill=lapse)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666") 
```

```{r}
t1 <- lapse_train %>%
  group_by(actuarial_age) %>%
  mutate(lapse = sum(lapse == "yes")) %>%
  select(actuarial_age, lapse) %>%
  distinct() %>%
  arrange(actuarial_age)

t2 <- lapse_train %>%
  group_by(actuarial_age) %>%
  mutate(exp = sum(exp)) %>%
  select(actuarial_age, exp) %>%
  distinct() %>%
  arrange(actuarial_age)

t <- t1 %>%
  bind_cols(exp = t2$exp, .name_repair = "unique") %>%
  mutate(`lapse/exp` = lapse/exp)

t %>%
  ggplot() +
  geom_point(aes(actuarial_age, `lapse/exp`), alpha = 0.4)

quantile(lapse_train$actuarial_age, probs = c(0.01, 0.99))
```


### exp

```{r}
table(ifelse(lapse_train$exp == 1, "=1", "<1"))/nrow(lapse_train)
lapse_train %>%
  ggplot(aes(x = exp)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666")
```

### duration
```{r}
lapse_train %>%
  ggplot(aes(x = duration, color=lapse, fill=lapse)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666") 

# Very significant

t1 <- lapse_train %>%
  mutate(duration = round(duration)) %>%
  group_by(duration) %>%
  mutate(lapse = sum(lapse == "yes")) %>%
  select(duration, lapse) %>%
  distinct() %>%
  arrange(duration)

t2 <- lapse_train %>%
  mutate(duration = round(duration)) %>%
  group_by(duration) %>%
  mutate(exp = sum(exp)) %>%
  select(duration, exp) %>%
  distinct() %>%
  arrange(duration)

t <- t1 %>%
  bind_cols(exp = t2$exp, .name_repair = "unique") %>%
  mutate(`lapse/exp` = lapse/exp)

t %>%
  ggplot() +
  geom_point(aes(duration, `lapse/exp`), alpha = 0.4)
```
### Capital
```{r}
summary(lapse_train$capital)
lapse_train %>%
  ggplot(aes(x = capital, color=lapse, fill=lapse)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666") 

lapse_train %>%
  filter(capital > 2000000)

lapse_train %>%
  ggplot(aes(x = capital, color=lapse, fill=lapse)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666") +
  scale_x_log10()
# here we have some outliers. We should take it into account.

t1 <- lapse_train %>%
  mutate(capital = round(capital)) %>%
  group_by(capital) %>%
  mutate(lapse = sum(lapse == "yes")) %>%
  select(capital, lapse) %>%
  distinct() %>%
  arrange(capital)

t2 <- lapse_train %>%
  mutate(capital = round(capital)) %>%
  group_by(capital) %>%
  mutate(exp = sum(exp)) %>%
  select(capital, exp) %>%
  distinct() %>%
  arrange(capital)

t <- t1 %>%
  bind_cols(exp = t2$exp, .name_repair = "unique") %>%
  mutate(`lapse/exp` = lapse/exp)

t %>%
  ggplot() +
  geom_point(aes(capital, `lapse/exp`), alpha = 0.4)
```


### IMC
```{r}
summary(lapse_train$IMC)

lapse_train %>%
  ggplot(aes(x = IMC, color=lapse, fill=lapse)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666") 

lapse_train %>%
  ggplot(aes(x = IMC, color=lapse, fill=lapse)) +
  geom_histogram(aes(y=..density..), alpha = 0.5)+
  geom_density(alpha=.2, fill="#FF6666") +
  scale_x_log10()

# Some outliers too
```

```{r}
names(lapse_train)

(a <- cor(lapse_train[, c(8, 10, 11, 13, 17)]))
corrplot::corrplot(a)

lapse_train %>%
  ggplot(aes(x = capital, y = IMC)) +
  geom_point()
# we have no linear correlation between the numerical variables
```















